### 2019-03-02 review/referee_response_0.txt

REFEREE COMMENT

2) As mentioned in 1), another possible interpretation for explaining the
deviation from a linear ephemeris of the TESS mid-transit times could be the
presence of anomalies in the light curves due to stellar activity. It has been
well established that WASP-4 is an active star with a rotation period (P) in
between 20-40 days (P ~ 22 days and P ~ 34 days were determined by
Sanchis-Ojeda et al. 2011 and Hoyer et al. 2013, respectively), which is
expected for a main-sequence G7 star. Given that the TESS mid-transits span
around 30 days, wouldn't it be possible that these measurements be affected by
the existence of unseen stellar spots? Regarding this point, I would like to
see an extra paragraph in the Section "Interpretation" with a discussion about
this scenario.

  > RESPONSE: We have considered this possibility, and have now added
  > the following text in Section 3.3:

  > To explore the effect of possible spot-crossing events on the TESS
  > transit time measurements, we performed a separate test.  We injected
  > triangular spot-anomalies with amplitude $0.03\%$ and duration 30
  > minutes at random phases into each transit.  The amplitudes were
  > larger than the spot-crossing anomalies observed by
  > \citet{southworth_high-precision_2009} and
  > \citet{sanchis-ojeda_starspots_2011}, and the durations were
  > comparable.  Spots of these amplitudes resemble the anomalies present
  > in transits ``1360.54'' and ``1372.58'' of Figure
  > \ref{fig:lightcurves}: they are visible, but at low confidence.
  > 
  > With spots injected, we repeated our measurement of the transit times.
  > On average, the measured transit times did not change with spots
  > injected, because the flux deviations are symmetric about the transit
  > minimum.  For individual transits, there were no cases for which the
  > timing deviation was larger than one minute.  The most concerning
  > shift occurs when the spot anomaly occurs during transit egress, in
  > which case the measured mid-time is shifted early by between 30 and 50
  > seconds \citep[qualitatively similar to results found
  > by][]{ioannidis_how_2016}.
  > 
  > Therefore the TESS observations could hypothetically all be skewed
  > early if there were spot-crossing events during every egress.  Two
  > arguments rule out this possibility.  (1) The lightcurve residuals do
  > not show evidence for these events.  (2) The stellar rotation period
  > is between 20 and 40 days, and the sky-projected stellar obliquity is
  > less than 10 degrees
  > \citep{triaud_spin-orbit_2010,sanchis-ojeda_starspots_2011,hoyer_tramos_2013}.
  > Since the planet orbits every 1.3 days, requiring that spot anomalies
  > always occur during egress would be equivalent to requiring a stellar
  > spot distribution that is exquisitely (and thus implausibly)
  > distributed to match the planet egress times.

@@

  > RESPONSE: Done, thank you.

REFEREE COMMENT

ii) How did the authors estimate the errors of the stellar and planet
parameters?

  > RESPONSE: We have added the following sentence in the last
  > paragraph of Section 2.3: "The uncertainties in our derived
  > stellar and planetary parameters are propagated according to
  > standard analytic formulae, under the assumption of uncorrelated
  > and normally distributed initial measurements."

@@

3. TIMING ANALYSIS

Section 3.1:

i) The sentence "..Since those data points carry significant weight in the
analysis, we checked that the timestamps in their data represent mid-exposure
times, that the barycentric correction was performed correctly, and that the
time system of the final results was BJD(TDB).." is confusing, because seems to
imply that these checks were only performed on the Huitson et al. mid-transit
times and not in all the literature mid-transit times. It would be
recommendable to clarify this point.

  > RESPONSE: we have modified the wording slightly to clarify. The
  > point is that for these Gemini South observations, we confirmed
  > with the authors through back and forth correspondence that the
  > details of their time-keeping were sound.  We made this extra
  > effort because of the weight of the Huitson+(2017) times in the
  > analysis.
  >
  > For the remaining data points, we used the references stated in
  > Section 3.1 and Table 2 to confirm that we were using with
  > mid-exposure times, and corrected from HJD_UTC to BJD_TDB as
  > appropriate. However the point is indeed that because of the
  > weight of the Huitson et al (2017) times, we made an particular
  > extra effort to confirm their accuracy with the original authors.

@@

REFEREE COMMENT

ii) Figure 3:

I would recommend changing the "title" indicated in bold font in the caption,
by a more explicit sentence about what is shown in the plot. To give an
example, a possibility could be "Observed minus predicted mid-transit times".

  > RESPONSE: Though we appreciate the suggestion, we have opted to
  > keep the "title" as is. The nuance of what "predicted" means for
  > this plot (a linear ephemeris, based only on the pre-TESS times)
  > is sufficient that we find it to be better left in the body of the
  > caption.

@@

REFEREE COMMENT

Also: ii-1) In the legends of both panels, could the authors specify which is
the bin size of the TESS light curves? (Please, fix this in all the
manuscript).

  > RESPONSE: We have added the following sentence in the caption:
  > "The binned TESS point is the weighted average of 18 TESS
  > transits".  The caption of Figure 4 already stated the same.  We
  > have changed the words "Dots" to "Symbols" in the caption of
  > Figure 6 in order to clarify this for the reader.

@@

  > RESPONSE: The caption now reads "Best-fit transit timing model
  > parameters".

REFEREE COMMENT

vii) Given that the mid-transit times values used in this work are taken from
different authors with different methodologies to perform the fitting and
detrending of the light curves, could the authors mention how this could affect
their results?

  > RESPONSE: We have added the following paragraphs at the end of
  > Section 3.3:

  > There is a final systematic concern in our global methodology.  We
  > have opted to use mid-transit time values derived by different
  > authors, who used heterogeneous methods to fit and detrend their
  > lightcurves.  This of course assumes that these authors have
  > correctly documented their time systems.  Further, though many
  > choices in transit-fitting ({\it e.g.}, parametrization of
  > limb-darkening and eccentricity) do not affect transit mid-time
  > measurements, different detrending approaches can asymmetrically
  > warp transits and shift mid-transit times.  The magnitude of this
  > systematic effect is hard to quantify, but the situation is fairly
  > clear from Figure~\ref{fig:times}.  Many independent authors
  > provided transit measurements shortly after WASP-4b's discovery,
  > so the data should on average be acceptable.
  > \citet{huitson_gemini_2017} provided the most important data from
  > epochs 0-1000, and so if their data were systematically affected
  > by detrending choices at the level of many times their reported
  > uncertainties, our timing anomaly would be suspect.  For this
  > reason, we paid careful attention to the
  > \citet{huitson_gemini_2017} data set, and corresponded with the
  > authors to confirm that their results are not affected by
  > systematic effects of the required amplitude.
  > 
  > None of the concerns mentioned in this subsection are likely
  > explanations for the observed timing variations. We proceed by
  > considering possible astrophysical explanations.

@@

  > RESPONSE: As noted in the caption of Figure 5, the decay timescale values
  > are obtained from Equation 15, assuming Qstar'=10^7.  We have slightly
  > modified the text for clarity on this point.

REFEREE COMMENT:

v) In the top panels of Figures 4 and 6, the orbital decay model evolution
shows that at the beginning the O-C values increase until, at some point, they
reach a maximum and after that, they start decreasing. What is the physical
explanation for that behavior? If the orbital period is diminishing, wouldn't
it be expected to see only a decreasing in the O-C values? Why the O-C values
would increase at these first epochs?

  > RESPONSE: To make Figure 4 and 6, we start with the observed
  > transit and occultation mid-times, as a function of epoch number.
  > We then fit a line to these data.  If the data were generated by a
  > constant period orbit (with gaussian uncertainties) the residuals
  > to such a fit would show no correlation with epoch number.
  > However if the data were generated by a planet with a constant
  > period derivative (with gaussian uncertainties), the residuals to
  > such a fit would always be a parabola.  This is because the linear
  > fit to the data will have contributions from the beginning,
  > middle, and end of the transit timeseries. Thus the best-fit
  > linear ephemeris (to data that actually have a decreasing period)
  > has a best-fit period somewhere between the minimum and maximum
  > period. At the earliest epochs, the best-fit period is an
  > underestimate, so you see the increase in O-C. At the latest
  > epochs, the best-fit period is an overestimate, so you see the
  > decrease.

@@

REFEREE COMMENT:

Section 4.2:

i) It might be informative to explicitly include the equation (14) of Ragozzine
& Wolf (2009) to see of which parameters is depending on.

  > RESPONSE: we have opted against this suggestion, chiefly to
  > encourage the interested reader to refer to the original study!

@@

REFEREE COMMENT:

ii) Similar to Figure 3, I would recommend changing the title indicated in bold
font in Figure 6, by a more explicit sentence about what is shown in the plot.
For eample "Different apsidal precession and orbital decay models for the O-C
values" or something like that.

  > RESPONSE: As in Figure 3, we appreciate the suggestion, but have
  > opted to keep the "title" as is. We have appended the sentence
  > "The two models may begin to diverge in the mid-2020s" as further
  > clarification for the reader.

@@

  > RESPONSE: We have added the phrase "similar to Jupiter". Section 4.2
  > already notes that our measured value for k_2p under the precession
  > hypothesis (Equation 18) is imprecise.

REFEREE COMMENT:

v) Would it be possible to introduce a definition of the reduced mass presented
in equation 21? Also, could the authors specify which values they used for L,
Mc, and Rc and provide references?

  > RESPONSE: We have clarified the origin of the values used when
  > evaluating Equation 21.  The tricky ones -- the mass and radius of
  > the convective envelope -- came from the MESA/MIST projects, for
  > which we have added appropriate citations.  Our thanks to the
  > referee for pointing out this omission.  We have opted to not
  > explicitly define the reduced mass, as it is standard knowledge in
  > the gravitational two-body problem.
